---
title: "Classification with Logistic Regression"
author: "Chris Bentz"
date: "14/10/2023"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Description
Logistic regression analyses of the feature vectors per character string (loaded from NaLaFi/results/features.csv). The results are stored in NaLaFi/results/LR. Note that the number of characters has to be chosen manually (via num.char = ``'').

# Load Packages
If the libraries are not installed yet, you need to install them using, for example, the command: install.packages("ggplot2"). For the Hrate package this is different, since it comes from github. The devtools library needs to be installed, and then the install_github() function is used.
```{r, message = F}
library(dplyr)
library(class)
library(gmodels)
library(caret)
```

# Load Data
Load data table with values per text file. 
```{r}
# load estimations from stringBase corpus
estimations.df <- read.csv("~/Github/NaLaFi/results/features.csv")
#head(features.csv)
```

Exclude subcorpora (if needed).
```{r}
#selected <- c("natural")
#estimations.df <- estimations.df[!(estimations.df$subcorpus %in% selected), ]
```

Split into separate files by length of chunks in characters.
```{r}
# choose number of characters
num.char = 1000
# subset data frame
estimations.df <- estimations.df[estimations.df$num.char == num.char, ]
```

Select relevant columns of the data frame, i.e. the measures to be included in classification and the ''corpus'' or ''subcorpus'' column.
```{r}
estimations.subset <- estimations.df[c("corpus",
                                       "huni.chars", 
                                       "hrate.chars", 
                                       "ttr.chars", 
                                       "rm.chars"
                                       )]
```

Remove NAs (whole row)
```{r}
estimations.subset <- na.omit(estimations.subset)
```

# Center and scale the data
```{r}
estimations.scaled <- cbind(estimations.subset[1], scale(estimations.subset[2:ncol(estimations.subset)]))
```

# Create Training and Test Sets
```{r}
# Generating seed
set.seed(1234)
# Randomly generating our training and test samples with a respective ratio of 2/3 and 1/3
datasample <- sample(2, nrow(estimations.scaled), replace = TRUE, prob = c(0.67, 0.33))
# Generate training set
train <- estimations.scaled[datasample == 1, 1:ncol(estimations.scaled)]
# Generate test set 
test <- estimations.scaled[datasample == 2, 1:ncol(estimations.scaled)]
```

# Building logistic regression model
The following code to run a logistic regression is adopted from https://datasciencedojo.com/blog/logistic-regression-in-r-tutorial/ (last accessed 16.01.2023).
```{r}
# logistic regression estimation of labels
log.model <- glm(as.factor(corpus) ~., data = train, family = binomial(link = "logit"))
summary(log.model)
# look at the training data and predicted values (y) to check the the dummy coding,
# i.e. "non-writing" is coded as 0, and "writing" as 1
head(train)
head(log.model$y)
```

# Prediction
Make predictions using the logistic regression model with "trained", i.e. estimated coefficients.
```{r}
log.predictions <- predict(log.model, test, type = "response")
head(log.predictions)
```
Assign a label according to the rule that the label is "writing" if the prediction probability is >0.5, else assign "non-writing". 
```{r}
log.prediction.rd <- ifelse(log.predictions > 0.5, "writing", "non-writing")
head(log.prediction.rd, 10)
```

# Model evaluation
```{r}
# creating a dataframe from known (true) test labels
test.labels <- data.frame(test$corpus)
# combining predicted and known classes
class.comparison <- data.frame(log.prediction.rd, test.labels)
# giving appropriate column names
names(class.comparison) <- c("predicted", "observed")
# inspecting our results table
head(class.comparison)
# get confusion matrix
cm <- confusionMatrix(as.factor(class.comparison$predicted), 
                      reference = as.factor(class.comparison$observed))
print(cm)
# get precision, recall, and f1 from the output list of confusionMatrix()
accuracy <- cm$overall['Accuracy']
f1 <- cm[["byClass"]]["F1"]
recall <- cm[["byClass"]]["Recall"]
precision <- cm[["byClass"]]["Precision"]
  
# prepare data frame with results
lr.results <- data.frame(accuracy, precision, recall, f1, row.names = NULL)
lr.results.rounded <- round(lr.results, 2)
print(lr.results.rounded)
```

Write to file.
```{r}
write.csv(lr.results.rounded, file = paste("~/Github/NaLaFi/results/LR/results_logReg_", 
                                      paste(num.char, ".csv", sep = ""), 
                                   sep = ""), row.names = F)
```