---
title: "Classification with linear SVM"
author: "Chris Bentz"
date: "17/01/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Packages
If the libraries are not installed yet, you need to install them using, for example, the command: install.packages("ggplot2"). For the Hrate package this is different, since it comes from github. The devtools library needs to be installed, and then the install_github() function is used.
```{r, message = F}
library(dplyr)
library(e1071)
library(class)
library(gmodels)
library(caret)
```

# Load Data
Load data table with values per text file. 
```{r}
# load estimations from stringBase corpus
estimations.df <- read.csv("~/Github/NaLaFi/results/features.csv")
#head(features.csv)
```

Exclude subcorpora (if needed).
```{r}
selected <- c("shuffled")
estimations.df <- estimations.df[!(estimations.df$subcorpus %in% selected), ]
```

Split into separate files by length of chunks in characters.
```{r}
# choose number of characters
num.char = 100
# subset data frame
estimations.df <- estimations.df[estimations.df$num.char == num.char, ]
```

Select relevant columns of the data frame, i.e. the measures to be included in classification and the ''corpus'' or ''subcorpus'' column.
```{r}
estimations.subset <- estimations.df[c("corpus",
                                       "huni.chars", 
                                       "hrate.chars", 
                                       "ttr.chars", 
                                       "rm.chars"
                                       )]
```

Remove NAs (whole row)
```{r}
estimations.subset <- na.omit(estimations.subset)
```

# Center and scale the data
```{r}
estimations.scaled <- cbind(estimations.subset[1], scale(estimations.subset[2:ncol(estimations.subset)]))
```

# Create Training and Test Sets
```{r}
# Generating seed
set.seed(1234)
# Randomly generating our training and test samples with a respective ratio of 2/3 and 1/3
datasample <- sample(2, nrow(estimations.scaled), replace = TRUE, prob = c(0.67, 0.33))
# Generate training set
train <- estimations.scaled[datasample == 1, 1:ncol(estimations.scaled)]
# Generate test set 
test <- estimations.scaled[datasample == 2, 1:ncol(estimations.scaled)]
```

# Building SVM
The following code to build a linear SVM is adopted from https://rpubs.com/cliex159/865583 (last accessed 17.01.2023).
```{r}
# svm classification
svm.model <- svm(corpus ~ ., 
                data = train, 
                type = "C-classification", 
                kernel = "linear", 
                scale = FALSE)
# list components of model
names(svm.model)
# check the levels
svm.model$levels
```

# Prediction
Make predictions using the svm model.
```{r}
svm.prediction <- predict(svm.model, test)
```

# Model evaluation
```{r}
# creating a dataframe from known (true) test labels
test.labels <- data.frame(test$corpus)
# combining predicted and known classes
class.comparison <- data.frame(svm.prediction, test.labels)
# giving appropriate column names
names(class.comparison) <- c("predicted", "observed")
# inspecting our results table
head(class.comparison)
# get confusion matrix
cm <- confusionMatrix(class.comparison$predicted, 
                      reference = class.comparison$observed)
print(cm)
# get precision, recall, and f1 from the output list of confusionMatrix()
f1 <- cm[["byClass"]]["F1"]
recall <- cm[["byClass"]]["Recall"]
precision <- cm[["byClass"]]["Precision"]
  
# prepare data frame with results
svm.results <- data.frame(precision, recall, f1, row.names = NULL)
svm.results.rounded <- round(svm.results, 2)
print(svm.results.rounded)
```

Write to file.
```{r}
write.csv(svm.results.rounded, file = paste("~/Github/NaLaFi/results/SVM/results_SVM_", 
                                      paste(num.char, ".csv", sep = ""), 
                                   sep = ""), row.names = F)
```

## Visualization
Build scatter plot of training dataset.
```{r}
huni.ttr.plot <- ggplot(data = train, aes(x = huni.chars, y = ttr.chars, color = corpus)) + 
    geom_point() 
huni.ttr.plot

w <- t(svm.model$coefs) %*% svm.model$SV
# calculate slope and intercept of decision boundary from weight vector and svm model
slope_1 <- -w[1]/w[2]
intercept_1 <- svm.model$rho/w[2]

#add decision boundary 
plot.decision <- huni.ttr.plot + geom_abline(slope = slope_1, intercept = intercept_1) 

#add margin boundaries
plot.margins <- plot.decision + 
 geom_abline(slope = slope_1, intercept = intercept_1 - 1/w[2], linetype = "dashed")+
 geom_abline(slope = slope_1, intercept = intercept_1 + 1/w[2], linetype = "dashed")

#display plot
plot.margins
```

Build scatter plot of training dataset.
```{r}
hrate.rm.plot <- ggplot(data = train, aes(x = hrate.chars, y = rm.chars, color = corpus)) + 
    geom_point() 
hrate.rm.plot

layered.plot <- 
    hrate.rm.plot + geom_point(data = train[svm.model$index, ], aes(x = hrate.chars, y = rm.chars), color = "purple", size = 4, alpha = 0.5)

#display plot
layered.plot
```