---
title: "Classification K-Nearest-Neighbors"
author: "Chris Bentz"
date: "14/10/2023"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Description
k-nearest neighbor analyses of the feature vectors per character string (loaded from NaLaFi/results/features.csv). The results are stored in NaLaFi/results/KNN. Note that the number of characters has to be chosen manually (via num.char = ""), likewise, the features to be included can be chosen in the lines defining "estimations.subset" below. Also, subcorpora can be excluded via the "selected" object. 

## Load Packages
If the libraries are not installed yet, you need to install them using, for example, the command: install.packages("ggplot2"). For the Hrate package this is different, since it comes from github. The devtools library needs to be installed, and then the install_github() function is used.
```{r, message = F}
library(ggplot2)
library(dplyr)
library(class)
library(gridExtra)
library(gmodels)
library(caret)
library(ggExtra)
library(ggpubr)
```

# Load Data
Load data table with values per text file. 
```{r}
# load estimations from stringBase corpus
estimations.df <- read.csv("~/Github/NaLaFi/results/features.csv")
#head(features.csv)
```

Exclude subcorpora (if needed). Choose the subcorpus to be excluded via "selected". 
```{r}
#selected <- c("udhr")
#estimations.df <- estimations.df[!(estimations.df$subcorpus %in% selected), ]
```

Split into separate files by length of chunks in characters.
```{r}
# choose number of characters
num.char = 100
# subset data frame
estimations.df <- estimations.df[estimations.df$num.char == num.char, ]
nrow(estimations.df)
```

Select relevant columns of the data frame, i.e. the measures to be included in classification and the ''corpus'' or ''subcorpus'' column.
```{r}
estimations.subset <- estimations.df[c("corpus", "subcorpus", 
                                       #"huni.chars", 
                                       #"hrate.chars", 
                                       #"ttr.chars", 
                                       "rm.chars"
                                       )]
```

Remove NAs (whole row)
```{r}
estimations.subset <- na.omit(estimations.subset)
```

# Center and scale the data
```{r}
estimations.scaled <- cbind(estimations.subset[1:2], scale(estimations.subset[3:ncol(estimations.subset)]))
nrow(estimations.scaled)
```

# Create Training and Test Sets
```{r}
# Generating seed
set.seed(1234)
# Randomly generating our training and test samples with a respective ratio of 2/3 and 1/3
datasample <- sample(2, nrow(estimations.scaled), replace = TRUE, prob = c(0.67, 0.33))
# Generate training set
estimations.training <- estimations.scaled[datasample == 1, 3:ncol(estimations.scaled)]
nrow(estimations.training)
# Generate test set 
estimations.test <- estimations.scaled[datasample == 2, 3:ncol(estimations.scaled)]
nrow(estimations.test)
```

# Get training and test labels
```{r}
# Generate training labels
training.labels <- estimations.scaled[datasample == 1, 1]
# Generate test labels
test.labels <- estimations.scaled[datasample == 2, 1]
```

# Initialize data frame
```{r}
knn.results <- data.frame(k = numeric(0), accuracy = numeric(0), precision = numeric(0), 
                             recall = numeric(0), f1 = numeric(0))
```

# Building knn classifier
```{r}
# choose maximum number of neighbors n
n = 10
# run a loop over different numbers of neighbors up to n
for (k in 1:n){
  # knn estimation of labels
  predictions.knn <- knn(train = as.data.frame(estimations.training), 
                         test = as.data.frame(estimations.test), 
                         cl = training.labels, k = k)
  
  # model evaluation
  # creating a dataframe from known (true) test labels
  test.labels <- data.frame(test.labels)
  # combining predicted and known classes
  class.comparison <- data.frame(predictions.knn, test.labels)
  # giving appropriate column names
  names(class.comparison) <- c("predicted", "observed")
  # inspecting our results table
  head(class.comparison)
  # get confusion matrix
  cm <- confusionMatrix(data = class.comparison$predicted, 
                      reference = as.factor(class.comparison$observed))
  # print(cm)
  # get precision, recall, and f1 from the output list of confusionMatrix()
  accuracy <- cm$overall['Accuracy']
  f1 <- cm[["byClass"]]["F1"]
  recall <- cm[["byClass"]]["Recall"]
  precision <- cm[["byClass"]]["Precision"]
  
  # prepare data frame with results
  local.results <- data.frame(k, accuracy, precision, recall, f1, row.names = NULL)
  local.results.rounded <- round(local.results, 2)
  # print(local.results.rounded)
  knn.results <- rbind(knn.results, local.results.rounded)
}
print(knn.results)
```

Write to file. Note that the file names have to be changed manually here.
```{r}
write.csv(knn.results, file = paste("~/Github/NaLaFi/results/KNN/knn_results_rm", 
                                    paste(num.char, ".csv", sep =""), 
                                    sep =""), row.names = F)
```
