---
title: "Classification K-Nearest-Neighbors"
author: "Chris Bentz"
date: "28/02/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Packages
If the libraries are not installed yet, you need to install them using, for example, the command: install.packages("ggplot2"). For the Hrate package this is different, since it comes from github. The devtools library needs to be installed, and then the install_github() function is used.
```{r, message = F}
library(ggplot2)
library(dplyr)
library(class)
library(gridExtra)
library(gmodels)
library(caret)
```

# Load Data
Load data table with quantitative measures per text file. 
```{r}
# load estimations from stringBase corpus
estimations.df.sb <- read.csv("/home/chris/Github/StringBase/code/Tables/output_stringBase.csv")
head(estimations.df.sb)
# load estimations from 100LC corpus
estimations.df.100lc <- read.csv("/home/chris/Github/StringBase/code/Tables/output_100LC_1Ksample.csv")
head(estimations.df.100lc)
```

Combine 100LC and stringBase estimations.
```{r}
# change labels in column ``subcorpus'' for 100LC (otherwise there are too many to plot)
estimations.df.100lc$subcorpus <- rep("writing", nrow(estimations.df.100lc))
estimations.df.combined <- rbind(estimations.df.100lc, estimations.df.sb)
```

Select subcorpora, these correspond to classes for the classifications.
```{r}
selected <- c("writing","ancient","paleolithic")
estimations.df.combined <- estimations.df.combined[estimations.df.combined$subcorpus %in% selected, ]
```

Select relevant columns of the data frame, i.e. the measures to be included in classification and the ``subcorpus'' column.
```{r}
estimations.subset <- estimations.df.combined[c("subcorpus", "huni.chars", "hrate.chars", "ttr.chars", "rm.chars")]
```

Remove NAs (whole row)
```{r}
estimations.subset <- na.omit(estimations.subset)
```

# Center and scale the data
```{r}
estimations.scaled <- cbind(estimations.subset[1], scale(estimations.subset[2:ncol(estimations.subset)]))
```

# Create Training and Test Sets
```{r}
# Generating seed
set.seed(1234)
# Randomly generating our training and test sampels with a respective ratio of 2/3 and 1/3
datasample <- sample(2, nrow(estimations.scaled), replace = TRUE, prob = c(0.5, 0.5))
# Generate training set
estimations.training <- estimations.scaled[datasample == 1, 2:ncol(estimations.scaled)]
# Generate test set 
estimations.test <- estimations.scaled[datasample == 2, 2:ncol(estimations.scaled)]
```

# Get training and test labels
```{r}
# Generate training labels
training.labels <- estimations.scaled[datasample == 1, 1]
# Generate test labels
test.labels <- estimations.scaled[datasample == 2, 1]
```

# Building knn classifier
```{r}
estimations.knn <- knn(train = estimations.training, test = estimations.test, cl = training.labels, k = 3)
```

# Model Evaluation
```{r}
# creating a dataframe from known (true) test labels
test.labels <- data.frame(test.labels)

# combining predicted and known species classes
class.comparison <- data.frame(estimations.knn, test.labels)

# giving appropriate column names
names(class.comparison) <- c("predicted", "observed")

# inspecting our results table
class.comparison

#get confusion matrix
cm <- confusionMatrix(class.comparison$predicted, reference = class.comparison$observed)

# get macro average of F1 score across classes
f1 <- cm[["byClass"]][ , "F1"]
f1.macro.avg <- sum(cm[["byClass"]][ , "F1"])/length(unique(class.comparison$observed))

# get macro average of recall score across classes
recall <- cm[["byClass"]][ , "Recall"]
recall.macro.avg <- sum(cm[["byClass"]][ , "Recall"])/length(unique(class.comparison$observed))

# get macro average of precision score across classes
precision <- cm[["byClass"]][ , "Precision"]
precision.macro.avg <- sum(cm[["byClass"]][ , "Precision"])/length(unique(class.comparison$observed))
```

# Prepare and safe data frame with results
```{r}
# combine in data frame
knn.results <- data.frame(precision, recall, f1)
macro.avgs <- c(precision.macro.avg, recall.macro.avg, f1.macro.avg)
knn.results <- rbind(knn.results, macro.avgs)
row.names(knn.results)[length(unique(class.comparison$observed)) + 1] <- "Macro avg." 
knn.results.rounded <- round(knn.results, 2)
print(knn.results.rounded)
```

Write to file.
```{r}
write.csv(knn.results.rounded, file = "/home/chris/Github/StringBase/code/Tables/knn_results.csv")
```