---
title: "Estimations of Quantitative Measures"
author: "Chris Bentz"
date: "05/01/2023"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries
If the libraries are not installed yet, you need to install them using, for example, the command: install.packages("ggplot2"). For the Hrate package this is different, since it comes from github. The devtools library needs to be installed, and then the install_github() function is used.
```{r}
library(stringr)
library(ggplot2)
library(ggrepel)
library(plyr)
library(entropy)
library(ggExtra)
library(gsubfn)
#library(devtools)
#install_github("dimalik/Hrate")
library(Hrate)
```

## List files
Create list with all the files in the directory "data". 
```{r}
# give file paths to the files to be processed
file.list <- list.files(path = "~/Github/NaLaFi/data", 
                        recursive = T, full.names = T)
head(file.list)
length(file.list)

#same for the teddi sample (downloaded from https://drive.switch.ch/index.php/s/MJv7xFkzqlzFn0y)
file.list.teddi <- list.files(path = "~/Data/TeDDi_dumps/Teddi_unifiedformat", 
                        recursive = T, full.names = T)
head(file.list.teddi)
length(file.list.teddi)

# downsample the number of teddi files
file.list.teddi <- sample(file.list.teddi, 1000)

#concatenate the two lists
file.list.combined <- c(file.list, file.list.teddi)
length(file.list.combined)
```

# Estimations per file
```{r}
# set counter
counter = 0
# set the maximal number of units (n) to be used for analysis
n = 1000
# initialize dataframe to append results to
estimations.df <- data.frame(filename = character(0), corpus = character(0), 
                             subcorpus = character(0), huni.chars = numeric (0), 
                             hrate.chars = numeric(0), ttr.chars = numeric(0), 
                             rm.chars = numeric(0)) 

# start time
start_time <- Sys.time()
for (file in file.list.combined) {
  try({ # if the processing failes for a certain file, there will be no output for this file,
  # but the try() function allows the loop to keep running  
  
  # basic processing
  # loading textfile
  textfile <- scan(file, what = "char", quote = "", comment.char = "", 
                   encoding = "UTF-8", sep = "\n" , skip = 0) 
  # remove the header lines beginning with '#'
  textfile <- textfile[!grepl('^#.*$', textfile)]
  # remove annotations marked by '<>'
  textfile <- gsub("<.*>","",textfile) 
  # print(head(textfile))
  # get filename
  filename <- basename(file) 
  # print(filename) # for visual inspection
  # get corpus category
  if (grepl('non-writing', file)){
    corpus <- 'non-writing'
  } else {
    corpus <- 'writing'
  }
  # get subcorpus category
  if (grepl('Teddi', file)){
    subcorpus <- 'teddi'
  } else {
    subcorpus <- sub("_.*", "", filename)
  }
  # print(subcorpus) # for visual inspection
  
  # Split into individual characters/signs 
  # remove tabs and parentheses, as well as star signs `*' and plus signs `+´
  # note that this might have to be tuned according to the text files included
  textfile <- str_replace_all(textfile, c("\\\t" = "", "\\(" = "", "\\)" = "", 
                                       "\\]" = "", "\\[" = "",  "\\}" = "", 
                                       "\\{" = "", "\\*" = "", "\\+" = ""))
  # split the textfile into individual utf-8 characters. Note that white spaces are
  # counted as utf-8 characters here.
  chars <- unlist(strsplit(textfile, ""))
  chars <- chars[1:n] # use only maximally n units
  chars <- chars[!is.na(chars)] # remove NAs for vectors which are already shorter than n
  # chars <- chars[chars != " "] # remove white spaces from character vector
  
  # unigram entropy estimation
  # calculate unigram entropy for characters
  chars.df <- as.data.frame(table(chars))
  # print(chars.df)
  huni.chars <- entropy(chars.df$Freq, method = "ML", unit = "log2")
  
  # entropy rate estimation
  # the values chosen for max.length and every.word will crucially 
  # impact processing time. In case of "max.length = NULL" the length is n units
  # here use the try() function since entropy rate estimation can fail when particular
  # special characters are in the strings
  hrate.chars <- try(get.estimate(text = chars, every.word = 1, max.length = NULL))
  if (class(hrate.chars) == "numeric") {
    hrate.chars <- hrate.chars
  } else {
    hrate.chars = "NA"
  }

  # calculate type-token ratio (ttr)
  ttr.chars <- nrow(chars.df)/sum(chars.df$Freq)
  
  # calculate repetition measure according to Sproat (2014)
  # the overall number of repetitions is the sum of frequency counts minus 1. 
  R <- sum(chars.df$Freq-1) 
  # calculate the number of adjacent repetitions
  r = 0
  if (length(chars) > 1){
    for (i in 1:(length(chars)-1)){
      if (chars[i] == chars[i+1]){
        r = r + 1
      } else {
        r = r + 0
      }
    }
    # calculate the repetition measure
    rm.chars <- r/R
  } else {
    rm.chars <- "NA"  
  }
  
  # append results to dataframe
  local.df <- data.frame(filename, corpus, subcorpus, huni.chars, 
                         hrate.chars, ttr.chars, rm.chars)
  estimations.df <- rbind(estimations.df, local.df)
  # counter
  counter <- counter + 1
  # print(counter)
  })
}
end_time <- Sys.time()
end_time - start_time
#estimations.df
```

# Safe outputs to file
```{r}
write.csv(estimations.df, paste("~/Github/NaLaFi/results/estimation", 
                                paste(n, "chars.csv", sep = ""), sep = ""), row.names = F)
```
