---
title: "Entropy Analyses for Characters and White-Space-Separated Strings"
author: "Chris Bentz"
date: "22/02/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries
If the libraries are not installed yet, you need to install them using, for example, the command: install.packages("ggplot2"). For the Hrate package this is different, since it comes from github. The devtools library needs to be installed, and then the install_github() function is used.
```{r}
library(stringr)
library(ggplot2)
library(ggrepel)
library(plyr)
library(ggExtra)
library(ggpubr)
```

# Load Data
Load data table with quantitative measures per text file. 
```{r}
# load estimations from stringBase corpus
estimations.df.sb <- read.csv("/home/chris/Github/StringBase/code/Tables/output_stringBase.csv")
head(estimations.df.sb)
# load estimations from 100LC corpus
estimations.df.100lc <- read.csv("/home/chris/Github/StringBase/code/Tables/output_100LC_1Ksample.csv")
head(estimations.df.100lc)
```

Add meta-information to 100LC (if needed).
```{r}
# load meta-info from 100LC
meta.info <- read.csv("/home/chris/Data/100LC_Dumps/csv/file.csv")
meta.info <- meta.info[, 1:12]
# merge with estimations
estimations.df.100lc.meta <- merge(estimations.df.100lc, meta.info, by = "filename")
```

Combine 100LC and stringBase estimations.
```{r}
# change labels in column ``subcorpus'' for 100LC (otherwise there are too many to plot)
estimations.df.100lc$subcorpus <- rep("writing", nrow(estimations.df.100lc))
estimations.df.combined <- rbind(estimations.df.100lc, estimations.df.sb)
```

Select subcorpora (if needed).
```{r}
selected <- c("writing", "ancient", "paleolithic")
estimations.df.combined <- estimations.df.combined[estimations.df.combined$subcorpus %in% selected, ]
```

# Scatterplots

## Entropy rate vs. unigram entropy for characters 
```{r, fig.width = 4, fig.height = 4, warning = FALSE}
# plot
huni.hrate.chars.plot <- ggplot(estimations.df.combined, 
                                aes(x = huni.chars, y = hrate.chars, 
                                    colour = subcorpus)) + 
  geom_point(alpha = 0.5, size  = 1) +
  # geom_smooth(method = "lm") +
  xlim(0, max(estimations.df.combined$huni.chars)) +
  ylim(0, max(estimations.df.combined$hrate.chars)) +
  #theme(legend.position = "bottom") +
  #geom_rug() +
  #geom_segment(x = 0, y = 0, xend = 10, yend = 10, colour = "black", 
            #linetype = "dashed", size = 0.3) +
  #geom_text(hjust = 0, nudge_x = 0, size = 2) +
  #geom_label_repel(aes(label = code), force = 0.5, force.pull = 5, label.size = 0.5, size = 3) +
  geom_label_repel(data = estimations.df.combined[estimations.df.combined$code == "sgr_0001" | 
                                                  estimations.df.combined$code == "vhc_0145" |
                                                  estimations.df.combined$code == "sum_0003" |  
                                                  estimations.df.combined$code == "cmn_0001" |
                                                  estimations.df.combined$code == "eng_0001" ,], 
                   label = c("Deer teeth", "Mammoth", "Sumerian", "Chinese", "English"), 
                   size = 3, arrow = arrow(length = unit(0.04, "npc"), 
                             type = "closed", ends = "last"), nudge_y = 0.7,
                             segment.size  = 0.3) +
  labs(x = "Unigram entropy for characters", y = "Entropy rate for characters") +
  theme(legend.position = "none")
huni.hrate.chars.plot <- ggMarginal(huni.hrate.chars.plot, groupFill = T, type = "histogram", colour = "white")
huni.hrate.chars.plot
```

## Safe complete figure to file
```{r}
ggsave("Figures/huni_hrate_chars.pdf", huni.hrate.chars.plot, dpi = 300, width = 4, height = 4, device = cairo_pdf)
```

## Unigram entropy vs repetition rate for characters 
```{r, fig.width = 12, fig.height = 8, warning = FALSE}
huni.rm.chars.plot <- ggplot(estimations.df.combined, 
                                  aes(x = huni.chars, y = rm.chars, colour = subcorpus)) + 
  geom_point(alpha = 0.8, size  = 1.5) +
  #geom_smooth(method = "lm") +
  xlim(0, max(estimations.df.combined$huni.chars)) +
  ylim(0, max(estimations.df.combined$rm.chars)) +
  #theme(legend.position = "bottom") +
  geom_rug() +
  #geom_text(hjust = 0, nudge_x = 0.1, size = 2) +
  geom_label_repel(aes(label = code), force = 0.5, force.pull = 5, label.size = 0.2, size = 3) +
  labs(x = "Unigram entropy for characters", y = "Repetition rate for characters") 
huni.rm.chars.plot
```

## Safe complete figure to file
```{r}
ggsave("Figures/huni_rm_chars.pdf", huni.rm.chars.plot, width = 13, height = 8, dpi = 300, 
       scale = 1, device = cairo_pdf)
```

## TTR vs. unigram entropy for characters
```{r, fig.width = 12, fig.height = 8, warning = FALSE}
huni.ttr.chars.plot <- ggplot(estimations.df.combined, 
                   aes(x = huni.chars, y = ttr.chars, 
                   colour = subcorpus)) + 
  geom_point(alpha = 0.8, size  = 1.5) +
  #theme(legend.position = "bottom") +
  geom_rug() +
  #geom_text(hjust = 0, nudge_x = 0.01, size = 2) +
  geom_label_repel(aes(label = code), force = 0.5, force.pull = 5, label.size = 0.2, size = 3) +
  labs(x = "Unigram entropy for characters", y = "TTR for characters") 
huni.ttr.chars.plot
```

## Safe complete figure to file
```{r, fig.width = 7, fig.height = 5}
ggsave("Figures/huni_ttr_chars.pdf", huni.ttr.chars.plot, width = 13, height = 8, dpi = 300,
       scale = 1, device = cairo_pdf)
```

## TTR vs. repetition rate for characters
```{r, fig.width = 6, fig.height = 4, warning = FALSE}
ttr.rm.chars.plot <- ggplot(estimations.df.combined, 
                   aes(x = ttr.chars, y = rm.chars, 
                   colour = subcorpus)) + 
  geom_point(alpha = 0.8, size  = 1) +
  theme(legend.position = "left") +
  #geom_rug() +
  #geom_text(hjust = 0, nudge_x = 0.01, size = 2) +
  #geom_label_repel(aes(label = code), force = 0.5, force.pull = 10, label.size = 0.2, size = 3) +
  geom_label_repel(data = estimations.df.combined[estimations.df.combined$code == "sgr_0001" | 
                                                  estimations.df.combined$code == "vhc_0145" |
                                                  estimations.df.combined$code == "sum_0003" |  
                                                  estimations.df.combined$code == "cmn_0001" |
                                                  estimations.df.combined$code == "eng_0001" ,], 
                   label = c("Deer teeth", "Mammoth", "Sumerian", "Chinese", "English"), 
                   size = 3, arrow = arrow(length = unit(0.03, "npc"), 
                             type = "closed", ends = "last"), nudge_y = 0.2,
                             segment.size  = 0.3) +
  labs(x = "TTR for characters", y = "Repetition rate for characters") 
ttr.rm.chars.plot <- ggMarginal(ttr.rm.chars.plot, groupFill = T, type = "histogram", colour = "white")
ttr.rm.chars.plot
```

## Safe complete figure to file
```{r, fig.width = 6, fig.height = 4}
ggsave("Figures/ttr_rm_chars.pdf", ttr.rm.chars.plot, width = 6, height = 4, dpi = 300,
       scale = 1, device = cairo_pdf)
```

## Combined Plots
```{r, fig.width = 10, fig.height = 4}
plots.combined <- ggarrange(huni.hrate.chars.plot, ttr.rm.chars.plot,
                    labels = c("a)", "b)"),
                    ncol = 2, nrow = 1, widths = c(1, 1.3) )
plots.combined
```

## Safe complete figure to file
```{r}
ggsave("Figures/plots_combined.pdf", plots.combined, width = 10, height = 4, dpi = 300,
       scale = 1, device = cairo_pdf)
```
