---
title: "Estimation Plots"
author: "Chris Bentz"
date: "03/01/2023"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries
If the libraries are not installed yet, you need to install them using, for example, the command: install.packages("ggplot2"). For the Hrate package this is different, since it comes from github. The devtools library needs to be installed, and then the install_github() function is used.
```{r}
library(stringr)
library(ggplot2)
library(ggrepel)
library(plyr)
library(ggExtra)
library(ggpubr)
```

# Load Data
Load data table with values per text file. 
```{r}
# load estimations from stringBase corpus
estimations10.df <- read.csv("~/Github/NaLaFi/results/estimation10chars.csv")
estimations100.df <- read.csv("~/Github/NaLaFi/results/estimation100chars.csv")
estimations1000.df <- read.csv("~/Github/NaLaFi/results/estimation1000chars.csv")
#head(estimations10.df.)
```

Select subcorpora (if needed).
```{r}
selected <- c("shuffled")
estimations10.df <- estimations10.df[!(estimations10.df$subcorpus %in% selected), ]
```

# Scatterplots

## Entropy rate vs. unigram entropy for characters 
```{r, fig.width = 4, fig.height = 4, warning = FALSE}
# plot
huni.hrate.chars.plot <- ggplot(estimations10.df, 
                                aes(x = huni.chars, y = hrate.chars, 
                                    colour = subcorpus)) + 
  geom_point(alpha = 0.5, size  = 1) +
  labs(x = "Unigram entropy for characters", y = "Entropy rate for characters") +
  theme(legend.position = "none")
huni.hrate.chars.plot <- ggMarginal(huni.hrate.chars.plot, groupFill = T, type = "histogram", colour = "white")
huni.hrate.chars.plot
```

## TTR vs. repetition rate for characters
```{r, fig.width = 6, fig.height = 4, warning = FALSE}
ttr.rm.chars.plot <- ggplot(estimations10.df, 
                   aes(x = ttr.chars, y = rm.chars, 
                   colour = subcorpus)) + 
  geom_point(alpha = 0.5, size  = 1) +
  theme(legend.position = "left") +
  labs(x = "TTR for characters", y = "Repetition rate for characters") 
ttr.rm.chars.plot <- ggMarginal(ttr.rm.chars.plot, groupFill = T, type = "histogram", colour = "white")
ttr.rm.chars.plot
```

## Combined Plots
```{r, fig.width = 10, fig.height = 4}
plots.combined <- ggarrange(huni.hrate.chars.plot, ttr.rm.chars.plot,
                    labels = c("a)", "b)"),
                    ncol = 2, nrow = 1, widths = c(1, 1.3) )
plots.combined
```

## Safe complete figure to file
```{r}
ggsave("~/Github/NaLaFi/figures/plots_combined.pdf", plots.combined, width = 10, height = 4, dpi = 300,
       scale = 1, device = cairo_pdf)
```
