---
title: "Hyperparameter Tuning"
author: "Chris Bentz"
date: "19/10/2023"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Packages
If the libraries are not installed yet, you need to install them using, for example, the command: install.packages("ggplot2"). For the Hrate package this is different, since it comes from github. The devtools library needs to be installed, and then the install_github() function is used.
```{r, message = F}
library(ggplot2)
library(gridExtra)
```

# Load Data
Load data table with values per text file. 
```{r}
# load results from folder
mlp.10chars.results <- read.csv("~/Github/NaLaFi/results/MLP/results_MLP_10chars_tanh_sse_rprop+.csv")
mlp.100chars.results <- read.csv("~/Github/NaLaFi/results/MLP/results_MLP_100chars_tanh_sse_rprop+.csv")
mlp.1000chars.results <- read.csv("~/Github/NaLaFi/results/MLP/results_MLP_1000chars_tanh_sse_rprop+.csv")
mlp.all <- rbind(mlp.10chars.results, mlp.100chars.results, mlp.1000chars.results)
# knn
knn.10.chars.results <- read.csv("~/Github/NaLaFi/results/KNN/knn_results_10.csv")
knn.10.chars.results$num.chars <- rep(10, nrow(knn.10.chars.results)) 
knn.100.chars.results <- read.csv("~/Github/NaLaFi/results/KNN/knn_results_100.csv")
knn.100.chars.results$num.chars <- rep(100, nrow(knn.100.chars.results)) 
knn.1000.chars.results <- read.csv("~/Github/NaLaFi/results/KNN/knn_results_1000.csv")
knn.1000.chars.results$num.chars <- rep(1000, nrow(knn.100.chars.results)) 
# bind all together
knn.all <- rbind(knn.10.chars.results, knn.100.chars.results, knn.1000.chars.results)
#head(knn.results)
```

# Hyperparameter Plots

# knn results
```{r, warning = F}
knn.plot <- ggplot(knn.all, aes(x = k, y = f1, 
                                color = as.factor(num.chars), 
                                group = as.factor(num.chars))) + 
    #geom_errorbar(aes(ymin=f1-sd(f1), ymax=f1+sd(f1)), width=.1) +
    geom_line() + 
    geom_point() +
    theme_minimal() +
    theme(legend.position = "bottom") +
    labs(color = 'Num. Chars.')
#knn.plot
```

# MLP results
Plot depth, i.e. number of hidden layers, versus F1 score.
```{r, warning = F}
mlp.plot.depth <- ggplot(mlp.all, aes(x = hidden.depth, y = f1,
                                          group = as.factor(num.char), 
                                          color = as.factor(num.char))) + 
    geom_point() +
    geom_smooth(method = "loess") +
    theme_minimal() +
    theme(legend.position = "bottom") +
    labs(color = 'Num. Chars.') 
#mlp.plot.depth
```

Plot number of hidden units versus F1 score.
```{r, warning = F}
mlp.plot.units <- ggplot(mlp.all, aes(x = hidden.size, y = f1,
                                          group = as.factor(num.char), 
                                          color = as.factor(num.char))) +  
    geom_point() +
    geom_smooth(method = "loess") +
    theme_minimal() +
    theme(legend.position = "bottom") +
    labs(color = 'Num. Chars.') 
```

# Combine plots
```{r, fig.width = 12, fig.height = 4, warning = F}
plots.combined <- grid.arrange(knn.plot, mlp.plot.depth, mlp.plot.units, ncol = 3)
print(plots.combined)
```

Safe to file.
```{r}
ggsave("~/Github/NaLaFi/figures/hyperParamPlot.pdf", plots.combined, width = 12, 
       height = 4, dpi = 300, scale = 1, device = cairo_pdf)
```

