---
title: "Sampler for Character Strings"
author: "Chris Bentz"
date: "26/06/2023"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries
If the libraries are not installed yet, you need to install them using, for example, the command: install.packages("ggplot2"). For the Hrate package this is different, since it comes from github. The devtools library needs to be installed, and then the install_github() function is used.
```{r}
library(stringr)
```

## List files
Create list with all the files in the directory "data". 
```{r}
# give file paths to the files to be processed
file.list <- list.files(path = "~/Github/NaLaFi/data", 
                        recursive = T, full.names = T)
head(file.list)
length(file.list)

#same for the teddi sample (downloaded from https://drive.switch.ch/index.php/s/MJv7xFkzqlzFn0y)
file.list.teddi <- list.files(path = "~/Data/TeDDi_dumps/Teddi_unifiedformat", 
                        recursive = T, full.names = T)
head(file.list.teddi)
length(file.list.teddi)

# downsample the number of teddi files 
# (using all 23K files would yield an extremely unbalanced sample)
file.list.teddi <- sample(file.list.teddi, 100)

#concatenate the two lists
file.list.combined <- c(file.list, file.list.teddi)
length(file.list.combined)
```

## Sampler for Character Strings
```{r}
# choose the length of chunks in number of characters
char.num <- 1000

# start time
start_time <- Sys.time()
for (file in file.list.combined){
  try({ # if the processing fails for a certain file, there will be no output for this file,
  # but the try() function allows the loop to keep running  
  
  # basic processing
  # loading textfile
  textfile <- scan(file, what = "char", quote = "", comment.char = "", 
                   encoding = "UTF-8", sep = "\n" , skip = 0) 
  # remove the header lines beginning with '#'
  textfile <- textfile[!grepl('^#.*$', textfile)]
  # remove annotations marked by '<>'
  textfile <- gsub("<.*>","",textfile) 
  # print(head(textfile))
  
  # Split into individual characters/signs 
  # remove tabs and parentheses, as well as star signs `*' and plus signs `+´
  # note that this might have to be tuned according to the text files included
  textfile <- str_replace_all(textfile, c("\\\t" = "", "\\(" = "", "\\)" = "", 
                                       "\\]" = "", "\\[" = "",  "\\}" = "", 
                                       "\\{" = "", "\\*" = "", "\\+" = ""))
  # split the textfile into individual utf-8 characters. Note that white spaces are
  # counted as utf-8 characters here and not removed (to remove them uncomment line below).
  chars <- unlist(strsplit(textfile, ""))
  # chars <- chars[chars != " "] # remove white spaces from character vector
  # split into list of chunks of size char.num (i.e. 10, 100, 1000)
  chunks.list <- split(chars, ceiling(seq_along(chars)/char.num))
  # remove chunks from list which are shorter than char.num (the last chunk likely is)
  chunks.list <- Filter(function(x) length(x) == char.num, chunks.list)
  # use "next" statement to exclude empty chunks list
  if (length(chunks.list) == 0) {
    next
  }
  # limit number of text chunks to a defined maximum (since some text files are much larger than others)
  max = 10
  chunks.list <- chunks.list[1:max]
  
  # prepare writing to file
  # get original filename
  filename <- basename(file)
  # create new file name 
  # note: this is dependent on the exact file extensions (!)
  if (grepl('non-writing', file)){
    new.filename <- paste('non-writing_', filename, sep = "")
  } else if (grepl('writing', file)){
    new.filename <- paste('writing_', filename, sep = "")
  } else {
    new.filename <- paste('writing_teddi_', filename, sep = "")
  }
  
  # write to file
  lapply(chunks.list, write, paste("~/Github/NaLaFi/samples/", 
                          paste(char.num, new.filename, sep = "_"), sep = ""), 
         append = TRUE, ncolumns = char.num, sep = "")
  })
}
end_time <- Sys.time()
end_time - start_time
```
