  ---
title: "Simple Corpus Stats"
author: "Chris Bentz"
date: "10/01/2023"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries
If the libraries are not installed yet, you need to install them using, for example, the command: install.packages("ggplot2").
```{r}
library(stringr)
library(ggplot2)
library(ggridges)
library(plyr)
```

## List files
List all the files in the directory "corpus". 
```{r}
file.list <- list.files(path = "~/Github/NaLaFi/data/", 
                        recursive = T, full.names = T)
#print(file.list)
length(file.list)
```
## Count number of files in writing and non-writing subfolders
Count how many files are in each of the highest level subfolders of "data", and create a dataframe with counts.

```{r}
#number of "animal" files
writing.count <- length(file.list[grepl("/writing", file.list)])
nonwriting.count <- length(file.list[grepl("non-writing", file.list)])

#create data frame
df <- data.frame(subfolder = c("writing", "non-writing"),
                 count = c(writing.count, nonwriting.count))
```

Create a bar plot with counts.
```{r, fig.height = 4, fig.width = 4}
counts.plot <- ggplot(data = df, aes(x = subfolder, y = count, fill = subfolder)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = count), vjust = -0.2, color = "black", size = 4) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45))
counts.plot
```

## Safe figure to file
```{r, fig.width = 4, fig.height = 4}
ggsave("~/Github/NaLaFi/figures/simpleStats_writing.pdf", counts.plot, dpi = 300, 
       scale = 1, device = cairo_pdf)
```

## Count number of files in subfolders
Count how many files are in each of the highest level subfolders of "data", and create a dataframe with counts.

```{r}
#number of "animal" files
animal.count <- length(file.list[grepl("animal", file.list)])
ancient.count <- length(file.list[grepl("ancient", file.list)])
heraldics.count <- length(file.list[grepl("heraldics", file.list)])
morse.count <- length(file.list[grepl("morse", file.list)])
natural.count <- length(file.list[grepl("natural", file.list)])
pycode.count <- length(file.list[grepl("pycode", file.list)])
random.count <- length(file.list[grepl("random", file.list)])
signlang.count <- length(file.list[grepl("signlang", file.list)])
shuffled.count <- length(file.list[grepl("shuffled", file.list)])
udhr.count <- length(file.list[grepl("udhr", file.list)])
weather.count <- length(file.list[grepl("weather", file.list)])

#create data frame
df <- data.frame(subfolder = c("animal", "ancient", "heraldics", "morse",
                               "natural", "pycode", "random", "signlang", 
                               "shuffled", "udhr", "weather"),
                 count = c(animal.count, ancient.count, heraldics.count, 
                           morse.count, natural.count, pycode.count, 
                           random.count, signlang.count, shuffled.count, 
                           udhr.count, weather.count))
```

Create a bar plot with counts.
```{r, fig.height = 4, fig.width = 6}
counts.plot <- ggplot(data = df, aes(x = subfolder, y = count, fill = subfolder)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = count), vjust = -0.2, color = "black", size = 4) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45))
counts.plot
```

## Safe figure to file
```{r, fig.width = 6, fig.height = 4.5}
ggsave("~/Github/NaLaFi/figures/simpleStats_counts.pdf", counts.plot, dpi = 300, 
       scale = 1, device = cairo_pdf)
```

# Lengths of files in characters

## Read files and count characters
```{r}
# set counter
counter = 0
# initialize dataframe to append results to
simpleStats.df <- data.frame(filename = character(0), subcorpus = character(0), 
                             num.lines = numeric(0), num.chars = numeric(0))

for (file in file.list) 
{
  # loading textfile
  textfile <- scan(file, what = "char", quote = "", comment.char = "", 
                   encoding = "UTF-8", sep = "\n", skip = 7)
  textfile <- gsub("\t","",textfile) # remove tabs
  textfile <- gsub("<.*>","",textfile) # remove annotations marked by '<>'
  #print(head(textfile))
  # get filename
  filename <- basename(file) 
  #print(filename) # for visual inspection
  # get subcorpus category
  subcorpus <- sub("_.*", "", filename)
  #print(subcorpus) # for visual inspection
  # count number of lines in text file
  num.lines <- length(textfile)
  # count the number of utf-8 characters in text file (note that this includes white 
  # spaces)
  num.chars <- sum(nchar(textfile, type ="chars"))
  #print(num.chars) # for visual inspection
  # append results to dataframe
  local.df <- data.frame(filename, subcorpus, num.lines, num.chars)
  simpleStats.df <- rbind(simpleStats.df, local.df)
  # counter
  counter <- counter + 1
  #print(counter)
}
simpleStats.df
```

# Density plot
Density plot for numbers of characters per data file and subcorpus.
```{r}
# select subcorpora (if applicable)
# selection <- c("writing", "shuffled")
#simpleStats.df <- simpleStats.df[simpleStats.df$subcorpus %in% selection, ]
# get means per subcorpus
mu <- ddply(simpleStats.df, "subcorpus", summarise, grp.mean = mean(log(num.chars)))
# plot densities with mean values as vertical lines
density.plot <- ggplot(simpleStats.df, aes(x = log(num.chars), y = subcorpus, fill = subcorpus)) +
  geom_density_ridges(alpha = 0.4) +
  theme_ridges() 
  #geom_density(alpha = 0.4) +
  #geom_vline(data = mu, aes(xintercept = grp.mean, color = subcorpus),
           #  linetype = "dashed")
  #facet_wrap(~subcorpus)
print(density.plot)
```

## Safe figure to file
```{r, fig.width = 8, fig.height = 4}
ggsave("~/Github/NaLaFi/figures/simpleStats_density.pdf", density.plot, dpi = 300, 
       scale = 1, device = cairo_pdf)
```



