---
title: "Stabilization Analyses for Characters"
author: "Chris Bentz"
date: "02/01/2023"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries
If the libraries are not installed yet, you need to install them using, for example, the command: install.packages("ggplot2"). For the Hrate package this is different, since it comes from github. The devtools library needs to be installed, and then the install_github() function is used.
```{r}
library(stringr)
library(ggplot2)
library(plyr)
library(entropy)
library(ggExtra)
library(gsubfn)
# library(devtools)
# install_github("dimalik/Hrate")
library(Hrate)
```

## List files
Create list with all the files in the directory "corpus". 
```{r}
# give file paths to the files to be processed
file.list <- list.files(path = "~/Github/NaLaFi/data/", 
                        recursive = T, full.names = T)
head(file.list)
length(file.list)
```

# Stabilization analysis per file
```{r}
# set counter
counter = 0
# set the maximal number of units (n), and the stepsize for stabilization analysis
# (i.e. in steps of how many units are values calculated?)
n = 1000
stepsize = 100
# initialize dataframe to append results to
stabilization.df <- data.frame(filename = character(0), subcorpus = character(0), 
                               code = character(0), huni.chars = numeric (0), 
                               hrate.chars = numeric(0), ttr = numeric(0), 
                               units = numeric(0)) 
# start time
start_time <- Sys.time()
for (file in file.list) 
{
  try({ # if the processing failes for a certain file, there will be no output for this file,
  # but the try() function allows the loop to keep running  
  
  # basic processing
  # loading textfile
  textfile <- scan(file, what = "char", quote = "", 
                   comment.char = "", encoding = "UTF-8", sep = "\n" , skip = 7, nmax = 20) # skip 7 first lines
  # remove annotations marked by '<>'
  textfile <- gsub("<.*>","",textfile) 
  # print(head(textfile))
  # get filename
  filename <- basename(file) 
  #print(filename) # for visual inspection
  # get subcorpus category
  subcorpus <- sub("_.*", "", filename)
  # print(subcorpus) # for visual inspection
  # get the three letter identification code + the running number
  code <- substring(substring(filename, regexpr("_", filename) + 1), 1, 8)
  
  # Split into individual characters/signs 
  # remove tabs and parentheses, as well as star signs `*' and plus signs `+´
  # note that this might have to be tuned according to the text files included
  textfile <- str_replace_all(textfile, c("\\\t" = "", "\\(" = "", "\\)" = "", 
                                       "\\]" = "", "\\[" = "",  "\\}" = "", 
                                       "\\{" = "", "\\*" = "", "\\+" = ""))
  # split the textfile into individual utf-8 characters. Note that white spaces are
  # counted as utf-8 characters here.
  chars <- unlist(strsplit(textfile, ""))
  chars <- chars[1:n] # use only maximally n units
  chars <- chars[!is.na(chars)] # remove NAs for vectors which are already shorter
  # than n
  # chars <- chars[chars != " "] # remove white spaces from character vector
  
  # run loop with stepsizes
  # define the number of units (i.e. characters) used for analyses (note that k is      
  # always either equal to or smaller than n)
  k = length(chars)
  for (i in 1:(k/stepsize))
  {  
    # unigram entropy estimation
    # calculate unigram entropy for characters
    chars.df <- as.data.frame(table(chars[1:(i*stepsize)]))
    # print(chars.df)
    huni.chars <- entropy(chars.df$Freq, method = "ML", unit = "log2")
    # entropy rate estimation
    # note: the values chosen for max.length and every.word will crucially 
    # impact processing time. max.length = NULL means all units in the file are     
    # considered.
    hrate.chars <- get.estimate(text = chars[1:(i*stepsize)], every.word = 1, 
                                max.length = NULL)  
    # calculate type-token ratio (ttr)
    ttr.chars <- nrow(chars.df)/sum(chars.df$Freq)
    # append results to dataframe
    local.df <- data.frame(filename, subcorpus, code, huni.chars, hrate.chars, 
                           ttr.chars, units = i*stepsize)
    stabilization.df <- rbind(stabilization.df, local.df)
  }
  # counter
  counter <- counter + 1
  # print(counter)
  })
}
end_time <- Sys.time()
end_time - start_time
# stabilization.df
```

# Stabilization plots

## Unigram entropy characters
```{r, fig.width = 20, fig.height = 20, warning = FALSE}
huni.chars.plot <- ggplot(stabilization.df, aes(x = units, y = huni.chars,
                                                colour = subcorpus)) + 
  geom_line(alpha = 0.8, size  = 1.5) +
  theme(legend.position = "bottom") +
  labs(x = "number of characters", y = "unigram entropy for characters") +
  facet_wrap(~code)
huni.chars.plot
```

## Safe figure to file
```{r, fig.width = 20, fig.height = 20}
ggsave("~/Github/NaLaFi/figures/stabilization_huni_chars.pdf", huni.chars.plot, dpi = 300, 
       scale = 1, device = cairo_pdf)
```

## Entropy rate characters
```{r, fig.width = 20, fig.height = 20, warning = FALSE}
hrate.chars.plot <- ggplot(stabilization.df, aes(x = units, y = hrate.chars,
                                                colour = subcorpus)) + 
  geom_line(alpha = 0.8, size  = 1.5) +
  theme(legend.position = "bottom") +
  labs(x = "number of characters", y = "entropy rate for characters") + 
  facet_wrap(~code)
hrate.chars.plot
```

## Safe figure to file
```{r, fig.width = 20, fig.height = 20}
ggsave("~/Github/NaLaFi/figures/stabilization_hrate_chars.pdf", hrate.chars.plot, dpi = 300, 
       scale = 1, device = cairo_pdf)
```

## TTR characters
```{r, fig.width = 20, fig.height = 20, warning = FALSE}
ttr.chars.plot <- ggplot(stabilization.df, aes(x = units, y = ttr.chars,
                                                colour = subcorpus)) + 
  geom_line(alpha = 0.8, size  = 1.5) +
  theme(legend.position = "bottom") +
  labs(x = "number of characters", y = "TTR for characters") +
  facet_wrap(~code)
ttr.chars.plot
```

## Safe figure to file
```{r, fig.width = 20, fig.height = 20}
ggsave("~/Github/NaLaFi/figures/stabilization_ttr_chars.pdf", ttr.chars.plot, dpi = 300, 
       scale = 1, device = cairo_pdf)
```